Multi-Instance Methods for Partially Supervised Image Segmentation

### 2013-09-24 partial_supervision/paper.tex
vlad first mail

We formulate weakly supervised multi-class image segmentation as a
multi-instance problem based on these candidate segments.  In multi-instance
learning~\citep{dietterich1997solving} each training example is given as a
multi-set of instances, called a bag.  Each instance is represented as a
feature vector $x$ and a label $y$.

A bag is labeled positive if it contains at least one positive example and
negative otherwise.  During training only the labels of the training bags, not
of the instances inside the bags, are known.  The goal is to learn a classifier
for unseen bags. 
Formally, let $\mathcal{X}$ be the set of instances. To simplify notation we
assume that bags are simply sets, not multi-sets.
\pagebreak\\
Then a bag is an element of
the power set $2^\mathcal{X}$ and the task is to learn a function
\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and
labels $y_i \in \{-1,+1\}$.  It is assumed that the $f_{MI}$ function stems
from a so-called underlying concept, given by an (unknown) function
$f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
\begin{equation}\eqlabel{multi_instance}
f_{MI}(X)= \max_{x \in X}\ f_{I}(x).
\end{equation}

Multi-instance learning is a natural formulation for image classification and
has been successfully applied to this task~\citep{zhou2007multi}. We propose to
go a step further and apply multi-instance learning to the task of object-class
segmentation in natural images by also classifying instances, not only bags.
In this we follow the work of \citet{liconvex2010} and \citet{zha2008joint}, who not
only learned $f_{MI}$, but also $f_{I}$.

In our model each image forms a bag, while the candidate segments correspond
to the instances contained in the bag. During learning only presence of object
classes is needed as bag-level supervision. By learning $f_{I}$, we are then
able to predict for individual segments whether they contain the object class
of interest, thereby obtaining a segmentation of the object.
To measure the performance of our algorithm we use a dataset that not only contains
image-level annotation, but also pixel-level annotation of object. This allows
us to judge the success of learning on instance level.

\section{Related Work}
\subsection{Object Segment Proposals}\label{related_segments}
Most work on multi-class segmentation focuses on strong supervision on
superpixel level. There is still little work on using candidate segments.  The
method we use for generating candidate segments is Constraint Parametric
Min-Cuts (CPMC) from \citet{carreira2010constrained}.  This method creates a
wide variety of overlapping segments. Support vector regression (SVR) is
trained on these segments to estimate the overlap of segments with ground truth
object-class labeling from the Pascal VOC dataset~\citep{pascal}. This provides
a ranking of candidate segments according to how ``object-like'' they are,
which allows for selecting only a limited number of very object-like segments.
The method performed well on a variety of datasets, building the basis of
a very successful entry to the Pascal VOC segmentation
challenge~\citep{li2010object}. A similar approach to whole-object segment
proposals was investigated by \citet{endres2010category}, but they did not compare
their results with the state-of-the-art approach of \citep{carreira2010constrained}.
%Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
%algorithm.

\subsection{Multi-Instance Methods}
Multi-instance learning was formally introduced in \citet{dietterich1997solving}.
Since then, many algorithms were proposed to solve the multi-instance learning problem
\citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
We discuss only those that are relevant to the present treatment.

@@

The framework described above yields an image-level and a segment-level
classification.  To obtain a pixel-level object-class segmentation, we have to
combine these.  Since we do not make use of the ground truth segmentation
during training, we cannot learn an optimal combination as in
\citet{li2010object} but perform a simple majority vote instead.  We merge
segments into pixel-level class labels by setting the label $y_x$ of a pixel
$x$ according to
\begin{equation}
    y_x = \argmax_{y \in Y} \#\{S_i | x \in S_i \land y_{S_i}=y \},
\end{equation}
where $Y$ is the set of possible object classes, $S_i$ enumerates all segments within an
image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
assigned the class with the highest number of class segments containing it.
This simple heuristic yields good results in practice.

>>>>

\enlargethispage{10mm}
We formulate weakly supervised multi-class image segmentation as a
multi-instance problem based on these candidate segments.  In multi-instance
learning~\citep{dietterich1997solving} each training example is given as a
multi-set of instances, called a bag.  Each instance is represented as a
feature vector $x$ and a label $y$.
A bag is labeled positive if it contains at least one positive example and
negative otherwise.  During training only the labels of the training bags, not
of the instances inside the bags, are known.  The goal is to learn a classifier
for unseen bags. 
Formally, let $\mathcal{X}$ be the set of instances. To simplify notation we
assume that bags are simply sets, not multi-sets.
\pagebreak\\
Then a bag is an element of
the power set $2^\mathcal{X}$ and the task is to learn a function
\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and
labels $y_i \in \{-1,+1\}$.  It is assumed that the $f_{MI}$ function stems
from a so-called underlying concept, given by an (unknown) function
$f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
\begin{equation}\eqlabel{multi_instance}
f_{MI}(X)= \max_{x \in X}\ f_{I}(x).
\end{equation}

Multi-instance learning is a natural formulation for image classification and
has been successfully applied to this task~\citep{zhou2007multi}. We propose to
go a step further and apply multi-instance learning to the task of object-class
segmentation in natural images by also classifying instances, not only bags.
In this we follow the work of \citet{liconvex2010} and \citet{zha2008joint}, who not
only learned $f_{MI}$, but also $f_{I}$.
In our model each image forms a bag, while the candidate segments correspond
to the instances contained in the bag. During learning only presence of object
classes is needed as bag-level supervision. By learning $f_{I}$, we are then
able to predict for individual segments whether they contain the object class
of interest, thereby obtaining a segmentation of the object.
To measure the performance of our algorithm we use a dataset that not only contains
image-level annotation, but also pixel-level annotation of object. This allows
us to judge the success of learning on instance level.

\section{Related Work}
\subsection{Object Segment Proposals}\label{related_segments}
Most work on multi-class segmentation focuses on strong supervision on
superpixel level. There is still little work on using candidate segments.  The
method we use for generating candidate segments is Constraint Parametric
Min-Cuts (CPMC) from \citet{carreira2010constrained}.  This method creates a
wide variety of overlapping segments. Support vector regression (SVR) is
trained on these segments to estimate the overlap of segments with ground truth
object-class labeling from the Pascal VOC dataset~\citep{pascal}. This provides
a ranking of candidate segments according to how ``object-like'' they are,
which allows for selecting only a limited number of very object-like segments.
The method performed well on a variety of datasets, building the basis of
a very successful entry to the Pascal VOC segmentation
challenge~\citep{li2010object}. A similar approach to whole-object segment
proposals was investigated by \citet{endres2010category}, but they did not compare
their results with the state-of-the-art approach of \citet{carreira2010constrained}.
%Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
%algorithm.

\subsection{Multi-Instance Methods}
Multi-instance learning was formally introduced by \citet{dietterich1997solving}.
Since then, many algorithms were proposed to solve the multi-instance learning problem
\citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
We discuss only those that are relevant to the present treatment.

@@

The framework described above yields an image-level and a segment-level
classification.  To obtain a pixel-level object-class segmentation, we have to
combine these.  Since we do not make use of the ground truth segmentation
during training, we cannot learn an optimal combination as
\citet{li2010object} did, but perform a simple majority vote instead.  We merge
segments into pixel-level class labels by setting the label $y_x$ of a pixel
$x$ according to:
\begin{equation}
    y_x = \argmax_{y \in Y} |\{S_i | x \in S_i \land y_{S_i}=y \}|,
\end{equation}
Here $Y$ is the set of possible object classes, $S_i$ enumerates all segments within an
image and $y_{S_i}$ is the label of segment $S_i$. In other words each pixel is
assigned the class with the highest number of class segments containing it.
This simple heuristic yields good results in practice.
